---
title: "ACST3006 Assignment"
author: "Toby Burgess"
date: '2023-03-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Introduction
### Question 1

In this question, we aim to analyze the optimal risky portfolio for two risky assets and study the relationship between the assets' correlation and the resulting possible portfolios. We will be using R to perform the calculations and visualize the results.

Firstly, using our portfolio means and sd, we can populate the table given using results given in the lectures.
```{r results='asis', echo=FALSE}
library(knitr)
library(kableExtra)
source("~/Uni/ACST3006/Assignment/Scripts/Q1.R", echo=FALSE)

formatted_table <- portfolio_stats_df %>%
  kable(format = "html", digits = 4, col.names = c("Weight Asset 1", "Weight Asset 2", "Portfolio Return", "Portfolio SD")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  column_spec(1, bold = F, border_right = T) %>%
  column_spec(2, bold = F, border_right = T) %>%
  column_spec(3, bold = F, border_right = T) %>%
  column_spec(4, bold = F)

print(formatted_table)
```
Now to find the optimal risky portfolio, 

\begin{equation}
E(\mathbf{R}) = \begin{bmatrix}
0.0015\\
0.008
\end{bmatrix}
\end{equation}

And:
\begin{align}
\text{Cov}(R_1, R_2) &= 0.4 \times 0.1 \times 0.05 \\
                     &= 0.002
\end{align}

We calculate the covariance matrix as:
\begin{equation}
V = \begin{bmatrix}
0.01 & 0.02 \\
0.02 & 0.025
\end{bmatrix}
\end{equation}

We get the inverse matrix as:

\begin{equation}
V^{-1} = \begin{bmatrix}
119.0476 & -95.2381 \\
-95.2381 & 476.1905
\end{bmatrix}
\end{equation}

So, we can find:

\begin{align}
A &= \begin{bmatrix}
0.0015\\
0.008
\end{bmatrix}'\begin{bmatrix}
119.0476 & -95.2381 \\
-95.2381 & 476.1905
\end{bmatrix}^{-1}\begin{bmatrix}
0.0015\\
0.008
\end{bmatrix} \\
&= 0.03440476 \\
B &= \begin{bmatrix}
0.0015\\
0.008
\end{bmatrix}'\begin{bmatrix}
119.0476 & -95.2381 \\
-95.2381 & 476.1905
\end{bmatrix}^{-1}\begin{bmatrix}
1\\
1
\end{bmatrix} \\
&= 3.404762 \\
C &= \begin{bmatrix}
1\\
1
\end{bmatrix}'\begin{bmatrix}
119.0476 & -95.2381 \\
-95.2381 & 476.1905
\end{bmatrix}^{-1}\begin{bmatrix}
1\\
1
\end{bmatrix}\\
&= 404.7619 \\
D &= \det\begin{bmatrix}
A & B \\
B & C
\end{bmatrix}
= AC - B^2 \\
&=2.333333
\end{align}

The optimal return can be then found with:

\begin{align}
E(R_c) &= \frac{B\,E(R_{p^*})-A}{C\,E(R_{p^*})-B}\\
E(R_{p^*}) &= \frac{B*0.004-A}{C*0.004-B} \\
&= 1.164\%
\end{align}

And the standard deviation:

\begin{align}
\sigma_{R_p^*} &= (\frac{C\,\{E(R_p^*)\}^2-2\,B\,E(R_p^*) +A}{D})^{\frac{1}{2}}\\
&=(\frac{C\,\{0.01164\}^2-2\,B\,\times0.01164 +A}{D})^{\frac{1}{2}} \\
&= 0.06540948
\end{align}


Now we find the capital allocation line:

\begin{align}
E(R_p) &= E(R_p^Z) +\frac{D\,\sigma_{R_{p^*}}}{C\,E(R_{p^*})-B}\times\sigma_{R_{p}} \\
E(R_p) &= 0.004 + 0.1168026\times\sigma_{R_{p}}
\end{align}

Plotting this, we get:

```{r, echo=FALSE}
q1plot1
```

We can also graph the set of possible portfolios when correlation is as follows:
```{r, echo=FALSE}
q1plot2
# Clear the environment ready for question 2
rm(list=ls())
```

### Question 2

We are given the monthly return for stocks, U.S. portfolios and country portfolios.

#### Part I
Let us calculate the mean return, standard deviation, correlations and variances/covariances on all available stocks / portfolios

```{r, echo=TRUE}
stock_returns <- fread("Data/Q2.csv")

# Manipulate Data ---------------------------------------------------------
# Store the investment categories and investment types
investment_categories <- as.character(unlist(stock_returns[1, ]))[-1]
investment_types <- as.character(unlist(stock_returns[2, ]))[-1]
stock_returns <- stock_returns[-c(1, 2)]
colnames(stock_returns) <- c("Date", as.vector(investment_types))

# Convert the data.table to long format using melt()
stock_returns_melt <- melt(stock_returns, id.vars = "Date", variable.name = "Investment", value.name = "Return")
stock_returns_melt[, Investment_Category := investment_categories[Investment], by = Investment]
stock_returns_melt <- stock_returns_melt[, .(Investment, Investment_Category, Date, Return)][, Return := as.numeric(Return)]

# Part I ------------------------------------------------------------------
# Calculate mean return
mean_returns <- stock_returns_melt[, .(Mean_Return = mean(Return)), by = .(Investment, Investment_Category)]
sd_returns <- stock_returns_melt[, .(SD_Return = sd(Return)), by = .(Investment, Investment_Category)]
risk_free_rate <- mean_returns[Investment == "US Riskfree", Mean_Return]

# Calculate the covariances and correlations among risky assets
risky_assets <- stock_returns_melt[Investment != "US Riskfree"]
risky_assets_wide <- dcast(risky_assets, Date ~ Investment, value.var = "Return")

# Remove the 'Date' column
risky_assets_wide[, Date:=NULL]

# Compute the covariance matrix
cov_matrix <- cov(risky_assets_wide)

# Compute the correlation matrix
cor_matrix <- cor(risky_assets_wide)

print(mean_returns)
print(sd_returns)
print(risk_free_rate)
print(cov_matrix)
print(cor_matrix)

```
#### Part II

To find the equation of the efficient frontier, we must:

1. Filter the data set for country portfolios
2. Calculate mean returns and covariance matrix
3. Apply the formulas for A, B, C and D as above
4. Apply the formula for the efficient frontier:

\begin{align}
Var(R_p) &= \frac{C\,\{E(R_p^*)\}^2-2\,B\,E(R_p^*) +A}{D}\\
&= 29.5754\,\{E(R_p^*)\}^2-0.499E(R_p^*) + 0.00038
\end{align}

We can do this in R with the following code:

```{r, echo=TRUE}
# Filter the data to keep only the country portfolios
country_portfolios <- risky_assets[Investment_Category == "Country Port"]

# Calculate the mean returns and covariance matrix for the country portfolios
mean_returns <- country_portfolios[, .(Mean_Return = mean(Return)), by = Investment]
cov_matrix <- cov(country_portfolios[, dcast(.SD, Date ~ Investment, value.var = "Return")][, Date:=NULL])

A <- t(mean_returns$Mean_Return) %*% solve(cov_matrix) %*% mean_returns$Mean_Return
B <- t(rep(1, 6)) %*% solve(cov_matrix) %*% mean_returns$Mean_Return
C <- t(rep(1, 6)) %*% solve(cov_matrix) %*% rep(1,6)
D <- A * C - B^2

# Define the function for the portfolio variance
portfolio_variance <- function(E_Rp) {
  return((C * E_Rp^2 - 2 * B * E_Rp + A) / D)
}

# Define the range of expected returns
E_Rp_range <- seq(min(mean_returns$Mean_Return), max(mean_returns$Mean_Return), length.out = 500)

# Calculate the portfolio variances
portfolio_variances <- sapply(E_Rp_range, portfolio_variance)

# Calculate the portfolio standard deviations (Sigmas)
portfolio_sigmas <- sqrt(portfolio_variances)

# Create a data frame with the expected returns and portfolio standard deviations
efficient_frontier_data <- data.frame(E_Rp = E_Rp_range, Sigma = portfolio_sigmas)

```

We can also find the Global Minimum Variance Portfolio, by applying the following R code:

```{r, echo=TRUE}
GMVP_Return <- B/C
GMVP_Var <- 1/C
GMVP_SD <- 1/sqrt(C)
```

```{r, echo=FALSE}
cat("The Global Minimum Variance Portfolio Return is: ", GMVP_Return)
cat("The Global Minimum Variance Portfolio Return is: ", GMVP_Var)
cat("The Global Minimum Variance Portfolio Return is: ", GMVP_SD)
```

This information can be displayed in a graph as below:

```{r, echo=FALSE}
q2plot1 <- ggplot() +
  geom_point(data = efficient_frontier_data, aes(x = Sigma, y = E_Rp), size = 0.5, color = "blue", alpha = 0.5) +
  geom_point(aes(x = GMVP_SD, y = GMVP_Return), color = "red", size = 4) +
  labs(x = "Portfolio Standard Deviation (Sigma)", y = "Portfolio Expected Return (E(R))",
       title = "Efficient Frontier and Global Minimum Variance Portfolio") +
  theme_minimal()

print(q2plot1)
```

### Part 3
We now consider an investor considering all country portfolio's, as well as the US risk-free portfolio.

We can calculate the equation for the CML by using the formula for optimal return and capital allocation line as above:

```{r, echo=TRUE}
optimal_return <- (A-risk_free_rate*B)/(B-risk_free_rate*C)
optimal_sd <- ((C*optimal_return^2-2*B*optimal_return+A)/D)^0.5

cml_slope <- (D*optimal_sd)/(C*optimal_return-B)
cml_intercept <- risk_free_rate
```

This gives us the following formula:
\begin{align}
E(R_p) &= 0.002955 + 0.2261123\times\sigma_{R_{p}}
\end{align}

Which can be graphed with the efficient frontier as below:
```{r, echo=FALSE}
cml_x <- seq(0, max(efficient_frontier_data$Sigma), 0.01)
cml_y <- cml_intercept + as.vector(cml_slope) * cml_x
cml_df <- data.frame(Return = cml_y, SD = cml_x)

# Graph CML
q2plot2 <- ggplot() +
  geom_line(data = cml_df, aes(x = SD, y = Return), color = "black", size = 1) +
  labs(x = "Portfolio Standard Deviation (Sigma)", y = "Portfolio Expected Return (E(R))",
       title = "Capital Market Line") +
  theme_minimal()

print(q2plot2)
```

##### Market Price of Risk
The market price of risk is defined as the slope of the Capital Market Line. In our case it is: `{r} cml_slope`
In practice it is the risk premium per extra unit of risk. It quantifies the compensation investors can expect to receive for taking on additional risk. A higher market price of risk indicates that investors are demanding a higher return for taking on additional risk. Conversely, a lower market price of risk suggests that investors are willing to accept a lower return for bearing the same level of risk.

##### Optimal Weights
We can find the optimal weights of this portfolio with the formula as follows:

\begin{align}
W^* = \begin{bmatrix}
w_1^* \\
w_2^*
  \end{bmatrix} &= V^{-1}
\begin{bmatrix}
E(\mathbf{R}), \mathbf{1} \end{bmatrix}\begin{bmatrix}
C & -B \\
-B & A
\end{bmatrix}\begin{bmatrix}
E(R_p*) \\
1
\end{bmatrix}\frac{1}{D}\\
\end{align}

In R:

```{r, echo=TRUE}
optimal_weights <- (solve(cov_matrix) %*% 
                      matrix(c(mean_returns$Mean_Return, rep(1,nrow(mean_returns))), nrow=nrow(mean_returns)) %*%
                      matrix(c(C, -B, -B, A), nrow=2) %*% 
                      matrix(c(optimal_return, 1), nrow = 2))/rep(D, 6)
```

```{r, echo=FALSE}
print(optimal_weights)
```

##### Portfolio Options
We can calculate the composition of various portfolios

```{r, echo=TRUE}
portfolio_options <- data.table(`Risky Portfolio` = c(0, 300000, 500000), `Risk Free Portfolio` = c(500000, 200000, 0))

portfolio_options[, `:=`(Return = `Risky Portfolio`*(1+optimal_return[1])+`Risk Free Portfolio`*(1+risk_free_rate), 
             SD = sqrt(`Risky Portfolio`^2*optimal_sd[1]^2))]
```


```{r results='asis', echo=FALSE}
portfolio_options <- portfolio_options[, lapply(.SD, function(x) scales::dollar(x))]

formatted_table <- portfolio_options %>%
  kable(format = "html", digits = 6, col.names = c("Risky Portfolio", "Risk Free Portfolio", "Return", "SD")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)

print(formatted_table)
```
##### Security Market Line and Beta's
We can calculate the Beta's for each portfolio as we know the mean returns ($R_i$), risk free rate ($R_f$) and market returns ($R_M$).

$$\beta_{iM}=\frac{E(R_i)-E(R_f)}{E(R_M)-E(R_f)}, $$

for each portfolio $i$


```{r, echo=TRUE}
mean_returns[, Beta:=(Mean_Return-risk_free_rate)/(optimal_return[1] - risk_free_rate)]

# Find sml line
sml_slope <- optimal_return - risk_free_rate
sml_intercept <- risk_free_rate
print(mean_returns)
```

We can plot this as:
```{r, echo=FALSE}
sml_x <- seq(0, 1.5, 0.01)
sml_y <- sml_intercept + as.vector(sml_slope) * sml_x

sml_df <- data.frame(Return = sml_y, SD = sml_x)

# Plot line
q2plot3 <- ggplot() +
  geom_line(data = sml_df, aes(x = SD, y = Return), color = "blue", size = 1) +
  labs(x = "Beta", y = "Portfolio Expected Return (E(R))",
       title = "Security Market Line") +
  theme_minimal()

print(q2plot3)

```

##### Interpretting Beta Values
The numerical beta values measure the systematic risk of an asset that cannot be diversified away. It is a measure of how the security / stock is correlated with the Market.
For example, Australia's $\beta_{iM} = 0.52726$. This means that when the market index (in this case the portfolio of countries) increases by 10%, the Australian portfolio will increase by 5.2726% and vice versa.

##### Example Portfolio
If we now consider a portfolio with a covariance of 0.0045 with the market, we get the following return:
```{r, echo=TRUE}
return = risk_free_rate + (0.0045/optimal_sd^2)*(optimal_return - risk_free_rate)

```

```{r, echo=FALSE}
cat("The return of this portfolio is:", return[1])
```





